Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-01 paper/HH4b.tex
work on MVA section
Work on the paper

\emailAdd{juan.rojo@physics.ox.ac.uk}

\preprint{OUTP-15-aaa}

\abstract{
A nice abstract
}

@@

% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sec-intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Monte Carlo samples
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sec-mcsamples}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% analysis strategy
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sec-analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Main results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sec-results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% MVA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sec-mva}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Conclusions and outlook
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{sec-conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\emailAdd{cigdem.issever@physics.ox.ac.uk}

\preprint{OUTP-15-14P}

\abstract{
  The measurement of Higgs pair production is one of the cornerstones of
  the LHC program for the next years.
  %
  Double Higgs production provides a unique window to the electroweak
  symmetry breaking mechanism, in particular to the Higgs trilinear coupling.
  %
  In this work we study the feasibility of the observation of the production of
  Higgs boson pairs at the High-Luminosity LHC using the
  $bb\bar{b}\bar{b}$ final state.
  %
  While the signal in this channel
  is enhanced by the large branching fraction of Higgs bosons to
  $b\bar{b}$ pairs, this measurement is  challenging due to the
  overwhelming QCD multijet background.
  %
  Our analysis is based on the combination of traditional cut-based
  methods and modern multivariate techniques, which lead to a substantial
  enhacement of the signal significance.
  %
  By combining the contribution from the boosted, intermediate and resolved
  topologies we achieve $S/\sqrt{B}=3.8$.
  %
  Therefore, we demonstrate that at the HL-LHC the $b\bar{b}b\bar{b}$ channel alone
  should allow to claim evidence of the observation of double Higgs production.
}

@@

\input{sec-intro}
\input{sec-mcsamples}
\input{sec-analysis}
\input{sec-results}
\input{sec-mva}
\input{sec-conclusions}

\appendix
\section{sec-appendix-overlap}

### 2015-07-01 paper/sec-analysis.tex
Work on the paper

>>>>

In section we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to the event topology.

{\bf needs flow chart}

First of all we discuss the settings and nomenclature that
we will use for jet clustering in this work.
%
Then we move to discuss how $b$-tagging is simulated,
following closely the ATLAS techniques.
%
Then we discuss the categorisation of events between different
topologies, and how we can prioritise among them.

\subsection{Jet reconstruction}

In our analysis we will use different types of jet definitions,
depending of the type of analysis that needs to be performed.
%
Therefore, we being by enumerating the choices that we make.
%
The starting point after the parton shower is the clustering
of all final state particles.
%
In this work all the jet reconstruction algorithms
are obtained from the {\tt FastJet} program~\cite{Cacciari:2011ma},
v3.1.0.
%
We adopt the following terminology for the different types of jets
that we will use in this work:

\begin{itemize}
\item {\it Small-$R$ jets}.

  These are jets  reconstructed with the
  anti-$k_T$ clustering algorithm~\cite{Cacciari:2008gp} with $R=0.4$ radius.
  %
  When reconstructing small-$R$ jets, we will keep only
  those with transverse momentum $p_T \ge 40$~GeV
  and pseudo-rapidity $|\eta|<2.5$ within the central
  acceptance of ATLAS and CMS, jets that fail to satisfy these
  conditions are discarded

  They are required to have $p_T > 50$~GeV and $|\eta|<2.5$.

\item {\it Large-$R$ jets}.

  These jets are also constructed with the
  anti-$k_T$ clustering algorithm, now using a $R=1.0$ radius.
  %
  In order to be accepted, these large-$R$ jets should
  satisfy  $p_T \ge 200$~GeV and lie in a pseudo-rapidity range by
  $|\eta|<2.0$.
  %
  If these conditions are not satisfied, the jet is discarded.
  %
  The more restrictive range  in pseudo-rapidity
  as compared to the small-$R$ jets,
  is motivated by mimicking the  experimental requirements
  in ATLAS and CMS
  related to the track-jet-based calibration.

  In addition to the basic $p_T$ and $\eta$
  selection requirements, large-$R$ jets should also
  satisfy the  BDRS mass-drop tagger~\cite{Butterworth:2008iy}
  conditions, in order
  to enhance the discrimination of the signal over the background
  events.
  %
  For the BDRS mass-drop tagger, we use the {\tt FastJet} default
  parameters of  $\mu = 0.67$ and $y_{\textrm{cut}}= 0.09$.

\item {\it Small-$R$ subjets}.

  Once we have reconstructed a suitable large-$R$ jet, important
  information can be obtained by looking at the kinematics of
  the small-$R$ subjets.
  %
  These subjets will be obtained by reclustering the constituents
  of the large-$R$ jet, always with the  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter $R=0.3$.
  %
  As we will discuss below, these small-$R$ subjets are the key of
  the $b$-tagging in the boosted category.
%
  \end{itemize}

\subsection{$b$-tagging}

Given the final state that needs to be reconstructed, the
optimisation of the $b$-tagging capabilities of the
LHC experiments is an essential requirement.
%
The $b$-tagging used in this feasibility study is inspired
in the current and projected ATLAS settings, and below
we also comment on the differences and similarities with the
corresponding CMS strategy.
%
For each type of jet defined above, a different
$b$-tagging strategy is used, as we discuss now.

\begin{itemize}

\item {\it Small-$R$ jets}.

  These are $b$-tagged as follows.
  %
  If a small-$R$ jet has at least one $b$-quark among their constituents,
  it will be tagged as a $b$-jet with probability $f_b$.
  %
  If no $b$-quarks are found among the constituents
  of this jet, it can be still be tagged as a $b$-jet with
  a mistag rate of $f_l$.
  %
  Note that the probability of tagging a jet is not increased
  if more than one $b$-quark is found among the jet constituents.

  \item {\it Large-$R$ jets}.

    Large-$R$ jets are $b$-tagged by ghost-associating anti-$k_T$ $R=0.3$
    subjets (defined as above) to the original large-$R$ jets.
    %
    In particular,
    a large-$R$ jet is considered  double-$b$-tagged if both
    the leading and subleading subjets (where the ordering
    is done in the subjet $p_T$) are both individually $b$-tagged.
    %
    Note that we only attempt to $b$-tag the two leading subjets:
    trying to $b$-tag all subjets above the $p_T$ cut degrades
    the signal significance due to the increase in combinatorics.

    Therefore, in this work we will exploit the capabilities of
    the LHC experiment of double-$b$-tagging inside large-$R$ jets.
    %
    Below we comment on the dependence of our results if single-$b$-tagging
    would be performed also for the large-$R$ jets, as done
    for the small-$R$ jets.

    Concerning the $b$-tagging probabilities and the
    light jet mistag probabilities, there are taken
    the same as for small-$R$ jets.
    %
    Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$, if only one of the two leading
    subjets has a $b$-quark then the tagging probability is
    $2f_bf_l$, and if none of the two have $b$-quarks
    as constituents, the mistag rate will be
    $f_l^2$.

\end{itemize}

Concerning the actual values of the $b$-tag probability $f_b$ and
the $b$-mistag probability of light jets $f_l$, in this work
we adopt as baseline the values $f_b=0.8$ and $f_l=0.01$.
%
In Sect.~\ref{sec:optimisation} we study the dependence of
our results against variations of $f_b$ and $f_l$ both
using more conservative and more aggressive values.
%
There we also study the dependence of the results
on other analysis settings related to the performance of
the experiments, for example by varying the momentum smearing.

\subsection{Event categorisation and cut flow}

In Table~\ref{sec:categorisation} we summarize the definitions of each
category

The di-Higgs system is reconstructed by considering all possibilities of forming two pairs of jets with invariant masses $m_{j1j2}$ and 
$m_{j3j4}$, respectively, and choosing the configuration that minimizes their difference $|m_{j1j2} - m_{j3j4}|$. Only the four leading-$p_T$ jets
are considered here.

The boosted selection requires the presence of at least two $bb$-tagged, selected anti-$k_T$ $R=1.0$ jets which are taken as the two Higgs boson candidates.

The intermediate topology requires the presence of at least one $bb$-tagged, selected anti-$k_T$ $R = 1.0$ jet. If there is more than one, the leading one is taken as the first Higgs boson candidate. In addition, there must be at least two anti-$k_T$ $R=0.4$ jets with a separation $\Delta R > 1.0$ with respect to the large-$R$ jet
that is selected as a Higgs boson candidate. The second Higgs boson candidate is reconstructed from the two small-$R$ jets that minimize the difference $|m_{fj} - m_{j1j2}|$.

The resolved selection requires the presence of at least four $b$-tagged, selected anti-$k_T$ $R=0.4$ jets.

All Higgs candidates are required to meet the following Higgs mass window cut: $|m_{H} - 125| < 40$~GeV

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{c|c|c|c}
    \hline
    &  Boosted  &  Intermediate  &  Resolved  \\
    \hline
    \hline 
    Jets  &  $\ge 2$ large-$R$ jets  & 1 large-$R$ jet, $\ge 2$ small-$R$
    jets  &  $\ge$ 4 small-$R$ jets \\
    Cuts on jets  & $p_T \ge 200$ GeV   &  $p_T \ge 200$ GeV (large-$R$ jet)
    &  $p_T \ge 40$ GeV \\
    &  & $p_T \ge 40$ GeV (small-$R$ jets)   &    \\
    &   $|\eta|<2.0$  & $|\eta|<2.0$ (large-$R$ jet) & $|\eta|<2.5$ \\
    &        & $|\eta|<2.5$ (small-$R$ jet)  &  \\
    \hline
    \end{tabular}
  \caption{\small Definition of the three event categories used in this
    work, together with the corresponding basic kinematical cuts.
\label{sec:categorisation}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-01 paper/sec-mcsamples.tex
more work on text
Work on the paper

\section{Monte Carlo samples}

Here we descrive the generation of the MC samples that
we are using in this work

>>>>

\section{Monte Carlo samples}
\label{mcgeneration}

First of all in this section we discuss the details of the
Monte Carlo generation of the signal and background
events used in this work.

\subsection{Signal}
%%%%%%%%%%%%%%%%%%%%%%%%%%

Higgs pair production is simulated at leading order using
{\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}, with finite $m_t$ effects taken into
account.\footnote{We thank Eleni Vrodynou for assistance here.}
%
We use the NNPDF 3.0 $n_f = 4$ LO set~\cite{Ball:2014uwa} with
$\alpha_S(m_Z^2)=0.118$
as interfaced via {\tt LHAPDF6}~\cite{Buckley:2014ana}.
%
We use the default settings of the renormalization and factorization
scale of the model.
%
The Higgs boson parameters are also the same in the default model,
in particular we use $m_H=125$ GeV, consistent with the latest
measurements from ATLAS and CMS.

%
Parton level events are then showered with the {\tt Pythia8} Monte
Carlo~\cite{Sjostrand:2007gs,Sjostrand:2014zea}, in particular with {\tt v8.201}.
%
We use the default settings in the parton shower, in particular
we use the Monash 2013 tune~\cite{Skands:2014pea}, based on the NNPDF2.3LO PDF set~\cite{Ball:2012cx}.
%
We have switched off hadronisation, since this allows a more transparent modeling of
$b$-tagging.
%
A complete simulation of $b$-tagging at the hadron level is difficult to achieve without a full
detector simulator, which is beyond the scope of this paper.
%
For a similar reason, we don't consider underlying event, multiple parton interactions and
pile-up, which can be taken of using a variety of subtraction methods~\cite{Cacciari:2009dp}.

The signal samples are rescaled so that the integrated cross-section matches that of the
NNLO calculation~\cite{deFlorian:2013jea}: this corresponds to using a $K$-factor $\sigma_{NNLO}/sigma_{LO}=2.3$.
%
Let us note that resummed NNLO+NNLL calculations for Higgs pair production have also become available recently~\cite{deFlorian:2015moa},
leading to a moderate enhancement as compared to the fixed-order NNLO calculation.

\subsection{Background}

Now we turn to discuss the Monte Carlo generation of the relevant background processes.
%
All background samples are generated with the {\tt SHERPA} event generator~\cite{Gleisberg:2008ta}, v2.1.1.
%
As in the case of the signal generation, the NNPDF 3.0 $n_f = 4$ LO set with strong coupling $\alpha_S=0.118$ is used for all samples.
Factorisation and renormalisation scales are set as $\mu_F=\mu_R=H_T/2$.

Samples!!!!

At the generation level the following basic cuts are applied.
%
Each final state particle in the hard process must have $p_T \ge 20$ GeV, and be located within $| \eta | \le 3.0$. All final state particles must be separated by a minimum $\Delta R_{\mathrm{min}} =0.1$.
%
We have checked that the generator-level cuts are loose enough as compared to the actual
analysis cuts.
%
Total cross-sections and details of the samples generated are shown in Table~\ref{tab:samples}.

%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Process &  Generator & $N_{\mathrm{evt}}$ & $\sigma_{\mathrm{tot}}$ \\
\hline
\hline
$pp \to HH$ &  {\tt MadGraph5\_aMC@NLO} & 100K & $1.729\times10^{-2}$ pb \\
\hline
\hline
$pp \to b\bar{b}b\bar{b}$ &  {\tt SHERPA}v2.1.1 & 3M &$1.121 \times10^3$ pb \\
$pp \to b\bar{b}jj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.659 \times 10^5$ pb \\
$pp \to jjjj$ &  {\tt SHERPA}v2.1.1 & 3M  & $9.709\times 10^6$ pb \\
$pp \to t\bar{t}$ &  {\tt SHERPA}v2.1.1 & 3M & $2.514\times 10^3$ pb \\
\hline
\end{tabular}
\caption{\small Summary of signal and background samples generated in this work,
together with the corresponding generator-level cross-section.} \label{tab:samples}
\end{center}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As a cross-check of the background cross-sections reported in Table~\ref{tab:samples}, we have produced multi-jet samples
using the {\tt MadGraph5\_aMC@NLO} program, and compared with the results for the same processes reported in
Ref.~\cite{Alwall:2014hca}.
%
For comparison with the latter numbers, 
we require in all samples four anti-$k_T$ $R=0.5$ jets with $p_T \ge 80 $ GeV, and the leading jet must have $p_T \ge 100$ GeV.
%
All jets must be within an acceptance of $|\eta| \le 2.5 $.
%
In the case of the samples with $b$ quarks in the final state, these requirements are extended to the appropriate number of $b$-jets. For example, in the 2$b$2$j$ sample there must be at least two $b$-jets that pass the cuts outlined above.
%
We find agreement, within the scale uncertainties, between the {\tt MadGraph5\_aMC@NLO} and {\tt SHERPA} calculations of the multi-jet
backgrounds, so we can be confident that the set-up that we will use in this analysis is robust enough.
